Data Exploration: IUCN and AquaMaps

### 2016-08-07 case_study_ohi/ohi_paired_maps.Rmd
push to Mazu to work remotely

library(ggplot2)
library(maps)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

@@

```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### get shapefile
### rasterize to half-degree grid; or raster::extract

if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {

  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
                       layer = 'ne_10m_bathymetry_K_200', 
                       stringsAsFactors = TRUE)

  ### LOICZID cell ID numbers - used in plotting rasters
  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  bath_list <- raster::extract(loiczid_raster, poly_bath, 
                             weights = TRUE, normalizeWeights = FALSE,
                             progress = 'text')
  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
    bind_rows() %>%
    rename(loiczid = value) %>%
    group_by(loiczid) %>%
    summarize(weight = sum(weight))

  ### generate raster?
  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
  values(bath_raster)[values(bath_raster) > 5] <- NA

  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
  ### probably want to filter to just 100% 200m+
}

@@

### rough recalcs
### assume that by removing the deep cells, we will not remove cells that are
### included in AquaMaps, since AquaMaps already filters by depth.  This means
### the overlap area will not be affected.
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#4dac26")  + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#b8e186") + 
  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#f1b6da") + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#d01c8b") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +

  geom_line(color = 'grey50', size = .5, alpha = .3) +
  geom_point(aes(color = method), size = 2, alpha = .8) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)')

>>>>

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

@@

```{r get_spp_cells, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read_csv(paired_maps_file) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps

# load IUCN cells, filter to just paired maps
spp_cells_iucn

# load AM cells, filter to just paired maps; include column of prob of occur

@@

### rough recalcs
### assume that by removing the deep cells, we will not remove cells that are
### included in AquaMaps, since AquaMaps already filters by depth.  This means
### the overlap area will not be affected.
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100,
           alpha = .3, fill= "#4dac26")  +
  annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100,
           alpha = .3, fill= "#b8e186") +
  annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0,
           alpha = .3, fill= "#f1b6da") +
  annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0,
           alpha = .3, fill= "#d01c8b") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +

  geom_line(color = 'grey50', size = .25, alpha = .3) +
  # geom_point(color = 'grey80', size = 2, alpha = 1) + ### blank behind points
  geom_point(aes(color = method), size = 2, alpha = .7) +
  scale_color_manual(values = c('original' = '#4d4dac', '200 m clip' = '#00cc00')) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)')

### 2016-08-07 data_explore/data_explore.Rmd
push to Mazu to work remotely

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc)) + #, 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b")

  ### plot the data points!
scatter_quadplot <- scatter_quadplot +
  geom_point(color = '#4d4dac', alpha = .6)

  ### label the quadrants:
scatter_quadplot <- scatter_quadplot +
  annotate("rect", xmin = 82, xmax = 98, ymin = 85, ymax = 91, 
           alpha = .7, fill= 'white')  + 
  annotate("rect", xmax = 24, xmin =  2, ymin = 85, ymax = 91, 
           alpha = .7, fill= "white") + 
  annotate("rect", xmin = 82, xmax = 98, ymax =  9, ymin = 15, 
           alpha = .7, fill= "white") + 
  annotate("rect", xmax = 20, xmin =  2, ymax =  9, ymin = 15, 
           alpha = .7, fill= "white") +
  annotate("text", x = 96, y = 88, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
           fontface = 'bold.italic', label = "Well-aligned") + 
  annotate("text", x =  4, y = 88, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Distribution-aligned") + 
  annotate("text", x = 96, y = 12, hjust = 1, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Area-aligned") + 
  annotate("text", x =  4, y = 12, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Poorly aligned")

>>>>

### define windows for labels
q1 <- c('x1' = 84, 'x2' = 98, 'y1' = 92.5, 'y2' = 97.5)
q2 <- c('x1' =  2, 'x2' = 22, 'y1' = 92.5, 'y2' = 97.5)
q3 <- c('x1' = 84, 'x2' = 98, 'y1' =  2.5, 'y2' = 7.5)
q4 <- c('x1' =  2, 'x2' = 18, 'y1' =  2.5, 'y2' = 7.5)

spp_list_labs <- spp_list %>%
  rename(x = area_ratio, y = sm_perc) %>%
  mutate(fade = FALSE,
         fade = ifelse((x > q1[1] & x < q1[2] & y > q1[3] & y < q1[4]), TRUE, fade),
         fade = ifelse((x > q2[1] & x < q2[2] & y > q2[3] & y < q2[4]), TRUE, fade),
         fade = ifelse((x > q3[1] & x < q3[2] & y > q3[3] & y < q3[4]), TRUE, fade),
         fade = ifelse((x > q4[1] & x < q4[2] & y > q4[3] & y < q4[4]), TRUE, fade))

### mongo plot time
scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b")

  ### plot the data points!
scatter_quadplot <- scatter_quadplot +
  # geom_point(color = 'white',   alpha = 1) + ### blank behind the points
  geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = .7))

  ### label the quadrants:
scatter_quadplot <- scatter_quadplot +
  annotate("text", x = 97, y = 95, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
           fontface = 'bold.italic', label = "Well-aligned") + 
  annotate("text", x =  3, y = 95, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Distribution-aligned") + 
  annotate("text", x = 97, y =  5, hjust = 1, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Area-aligned") + 
  annotate("text", x =  3, y =  5, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Poorly aligned")
